Automated Residual Plot Assessment with the R Package autovi and the Shiny App autovi.web

### 2025-05-09 autovi_paper.qmd
Give better name for the extended class

```{r echo = TRUE}
AUTO_GLM_VI <- new_class(AUTO_VI, class_name = "AUTO_GLM_VI")
register_method(
  AUTO_GLM_VI,
  get_fitted_and_resid = function(fitted_model = self$fitted_model) {
    tibble(.fitted = fitted(fitted_model),
           .resid = resid(fitted_model, type = "pearson"))
  },
  null_method = function(fitted_model = self$fitted_model) {
    dat <- model.frame(fitted_model)
    dat[[1]] <- rpois(nrow(dat), lambda = fitted(fitted_model))
    new_mod <- update(fitted_model, data = dat)
    return(self$get_fitted_and_resid(new_mod))
  }
)
auto_glm_vi <- AUTO_GLM_VI$instantiate
```

@@

```r
set.seed(10086)
x <- rnorm(300, sd = 0.5)
y <- rpois(300, lambda = exp(1 + x + x^2))
y_fit <- glm(y ~ x, family = "poisson")

glm_checker <- auto_glm_vi(glm(y ~ x, family = "poisson"),
                           keras_model = get_keras_model("vss_phn_32"))
glm_checker$plot_lineup()
```

```{r include = FALSE}
set.seed(10086)
x <- rnorm(300, sd = 0.5)
y <- rpois(300, lambda = exp(1 + x + x^2))
y_fit <- glm(y ~ x, family = "poisson")

glm_checker <- auto_glm_vi(glm(y ~ x, family = "poisson"))
```

```{r}
glm_checker$plot_lineup()
```

```r
glm_checker$check()
```

```{r message = TRUE}
if (!file.exists("cached_data/autovi_glm_checker_demo_print.rds")) {
  glm_checker <- auto_glm_vi(glm(y ~ x, family = "poisson"),
                           keras_model = get_keras_model("vss_phn_32"))
  glm_checker$check()
  saveRDS(glm_checker$..str..(), "cached_data/autovi_glm_checker_demo_print.rds")
} else {
  cli::cli_h3(readRDS("cached_data/autovi_glm_checker_demo_print.rds"))
}
```

>>>>

```{r echo = TRUE}
AUTO_POIS_VI <- new_class(AUTO_VI, class_name = "AUTO_POIS_VI")
register_method(
  AUTO_POIS_VI,
  get_fitted_and_resid = function(fitted_model = self$fitted_model) {
    tibble(.fitted = fitted(fitted_model),
           .resid = resid(fitted_model, type = "pearson"))
  },
  null_method = function(fitted_model = self$fitted_model) {
    dat <- model.frame(fitted_model)
    dat[[1]] <- rpois(nrow(dat), lambda = fitted(fitted_model))
    new_mod <- update(fitted_model, data = dat)
    return(self$get_fitted_and_resid(new_mod))
  }
)
auto_pois_vi <- AUTO_POIS_VI$instantiate
```

@@

```r
x <- rnorm(300, sd = 0.5)
y <- rpois(300, lambda = exp(1 + x + x^2))
pois_checker <- auto_pois_vi(
  glm(y ~ x, family = "poisson"),
  keras_model = get_keras_model("vss_phn_32")
)
pois_checker$plot_lineup()
```

```{r include = FALSE}
set.seed(10086)
x <- rnorm(300, sd = 0.5)
y <- rpois(300, lambda = exp(1 + x + x^2))
pois_checker <- auto_pois_vi(glm(y ~ x, family = "poisson"))
```

```{r}
pois_checker$plot_lineup()
```

```r
pois_checker$check()
```

```{r message = TRUE}
if (!file.exists("cached_data/autovi_glm_checker_demo_print.rds")) {
  pois_checker <- auto_pois_vi(glm(y ~ x, family = "poisson"),
                               keras_model = get_keras_model("vss_phn_32"))
  pois_checker$check()
  saveRDS(pois_checker$..str..(), "cached_data/autovi_glm_checker_demo_print.rds")
} else {
  cli::cli_h3(readRDS("cached_data/autovi_glm_checker_demo_print.rds"))
}
```

### 2025-05-09 autovi_paper.tex
Give better name for the extended class

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{AUTO\_GLM\_VI }\OtherTok{\textless{}{-}} \FunctionTok{new\_class}\NormalTok{(AUTO\_VI, }\AttributeTok{class\_name =} \StringTok{"AUTO\_GLM\_VI"}\NormalTok{)}
\FunctionTok{register\_method}\NormalTok{(}
\NormalTok{  AUTO\_GLM\_VI,}
  \AttributeTok{get\_fitted\_and\_resid =} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{fitted\_model =}\NormalTok{ self}\SpecialCharTok{$}\NormalTok{fitted\_model) \{}
    \FunctionTok{tibble}\NormalTok{(}\AttributeTok{.fitted =} \FunctionTok{fitted}\NormalTok{(fitted\_model),}
           \AttributeTok{.resid =} \FunctionTok{resid}\NormalTok{(fitted\_model, }\AttributeTok{type =} \StringTok{"pearson"}\NormalTok{))}
\NormalTok{  \},}
  \AttributeTok{null\_method =} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{fitted\_model =}\NormalTok{ self}\SpecialCharTok{$}\NormalTok{fitted\_model) \{}
\NormalTok{    dat }\OtherTok{\textless{}{-}} \FunctionTok{model.frame}\NormalTok{(fitted\_model)}
\NormalTok{    dat[[}\DecValTok{1}\NormalTok{]] }\OtherTok{\textless{}{-}} \FunctionTok{rpois}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(dat), }\AttributeTok{lambda =} \FunctionTok{fitted}\NormalTok{(fitted\_model))}
\NormalTok{    new\_mod }\OtherTok{\textless{}{-}} \FunctionTok{update}\NormalTok{(fitted\_model, }\AttributeTok{data =}\NormalTok{ dat)}
    \FunctionTok{return}\NormalTok{(self}\SpecialCharTok{$}\FunctionTok{get\_fitted\_and\_resid}\NormalTok{(new\_mod))}
\NormalTok{  \}}
\NormalTok{)}
\NormalTok{auto\_glm\_vi }\OtherTok{\textless{}{-}}\NormalTok{ AUTO\_GLM\_VI}\SpecialCharTok{$}\NormalTok{instantiate}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{10086}\NormalTok{)}
\NormalTok{x }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{300}\NormalTok{, }\AttributeTok{sd =} \FloatTok{0.5}\NormalTok{)}
\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{rpois}\NormalTok{(}\DecValTok{300}\NormalTok{, }\AttributeTok{lambda =} \FunctionTok{exp}\NormalTok{(}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ x }\SpecialCharTok{+}\NormalTok{ x}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{))}
\NormalTok{y\_fit }\OtherTok{\textless{}{-}} \FunctionTok{glm}\NormalTok{(y }\SpecialCharTok{\textasciitilde{}}\NormalTok{ x, }\AttributeTok{family =} \StringTok{"poisson"}\NormalTok{)}

\NormalTok{glm\_checker }\OtherTok{\textless{}{-}} \FunctionTok{auto\_glm\_vi}\NormalTok{(}\FunctionTok{glm}\NormalTok{(y }\SpecialCharTok{\textasciitilde{}}\NormalTok{ x, }\AttributeTok{family =} \StringTok{"poisson"}\NormalTok{),}
                           \AttributeTok{keras\_model =} \FunctionTok{get\_keras\_model}\NormalTok{(}\StringTok{"vss\_phn\_32"}\NormalTok{))}
\NormalTok{glm\_checker}\SpecialCharTok{$}\FunctionTok{plot\_lineup}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{glm\_checker}\SpecialCharTok{$}\FunctionTok{check}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
-- <AUTO_GLM_VI object>
Status:
 - Fitted model: glm, lm
 - Keras model: (None, 32, 32, 3) + (None, 5) -> (None, 1)
    - Output node index: 1
 - Result:
    - Observed visual signal strength: 4.875 (p-value = 0.009901)
    - Null visual signal strength: [100 draws]
       - Mean: 1.331
       - Quantiles: 
          ╔══════════════════════════════════════════╗
          ║  25%   50%   75%   80%   90%   95%   99% ║
          ║1.035 1.233 1.488 1.644 1.941 2.276 2.639 ║
          ╚══════════════════════════════════════════╝
    - Bootstrapped visual signal strength: [100 draws]
       - Mean: 5.51 (p-value = 0.009901)
       - Quantiles: 
          ╔══════════════════════════════════════════╗
          ║  25%   50%   75%   80%   90%   95%   99% ║
          ║5.330 5.505 5.698 5.735 5.830 5.903 6.013 ║
          ╚══════════════════════════════════════════╝
    - Likelihood ratio: 0.05096 (boot) / 0 (null) = Extremely large 
\end{verbatim}

>>>>

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{AUTO\_POIS\_VI }\OtherTok{\textless{}{-}} \FunctionTok{new\_class}\NormalTok{(AUTO\_VI, }\AttributeTok{class\_name =} \StringTok{"AUTO\_POIS\_VI"}\NormalTok{)}
\FunctionTok{register\_method}\NormalTok{(}
\NormalTok{  AUTO\_POIS\_VI,}
  \AttributeTok{get\_fitted\_and\_resid =} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{fitted\_model =}\NormalTok{ self}\SpecialCharTok{$}\NormalTok{fitted\_model) \{}
    \FunctionTok{tibble}\NormalTok{(}\AttributeTok{.fitted =} \FunctionTok{fitted}\NormalTok{(fitted\_model),}
           \AttributeTok{.resid =} \FunctionTok{resid}\NormalTok{(fitted\_model, }\AttributeTok{type =} \StringTok{"pearson"}\NormalTok{))}
\NormalTok{  \},}
  \AttributeTok{null\_method =} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{fitted\_model =}\NormalTok{ self}\SpecialCharTok{$}\NormalTok{fitted\_model) \{}
\NormalTok{    dat }\OtherTok{\textless{}{-}} \FunctionTok{model.frame}\NormalTok{(fitted\_model)}
\NormalTok{    dat[[}\DecValTok{1}\NormalTok{]] }\OtherTok{\textless{}{-}} \FunctionTok{rpois}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(dat), }\AttributeTok{lambda =} \FunctionTok{fitted}\NormalTok{(fitted\_model))}
\NormalTok{    new\_mod }\OtherTok{\textless{}{-}} \FunctionTok{update}\NormalTok{(fitted\_model, }\AttributeTok{data =}\NormalTok{ dat)}
    \FunctionTok{return}\NormalTok{(self}\SpecialCharTok{$}\FunctionTok{get\_fitted\_and\_resid}\NormalTok{(new\_mod))}
\NormalTok{  \}}
\NormalTok{)}
\NormalTok{auto\_pois\_vi }\OtherTok{\textless{}{-}}\NormalTok{ AUTO\_POIS\_VI}\SpecialCharTok{$}\NormalTok{instantiate}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{300}\NormalTok{, }\AttributeTok{sd =} \FloatTok{0.5}\NormalTok{)}
\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{rpois}\NormalTok{(}\DecValTok{300}\NormalTok{, }\AttributeTok{lambda =} \FunctionTok{exp}\NormalTok{(}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ x }\SpecialCharTok{+}\NormalTok{ x}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{))}
\NormalTok{pois\_checker }\OtherTok{\textless{}{-}} \FunctionTok{auto\_pois\_vi}\NormalTok{(}
  \FunctionTok{glm}\NormalTok{(y }\SpecialCharTok{\textasciitilde{}}\NormalTok{ x, }\AttributeTok{family =} \StringTok{"poisson"}\NormalTok{),}
  \AttributeTok{keras\_model =} \FunctionTok{get\_keras\_model}\NormalTok{(}\StringTok{"vss\_phn\_32"}\NormalTok{)}
\NormalTok{)}
\NormalTok{pois\_checker}\SpecialCharTok{$}\FunctionTok{plot\_lineup}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pois\_checker}\SpecialCharTok{$}\FunctionTok{check}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
-- <AUTO_POIS_VI object>
Status:
 - Fitted model: glm, lm
 - Keras model: (None, 32, 32, 3) + (None, 5) -> (None, 1)
    - Output node index: 1
 - Result:
    - Observed visual signal strength: 4.875 (p-value = 0.009901)
    - Null visual signal strength: [100 draws]
       - Mean: 1.331
       - Quantiles: 
          ╔══════════════════════════════════════════╗
          ║  25%   50%   75%   80%   90%   95%   99% ║
          ║1.035 1.233 1.488 1.644 1.941 2.276 2.639 ║
          ╚══════════════════════════════════════════╝
    - Bootstrapped visual signal strength: [100 draws]
       - Mean: 5.51 (p-value = 0.009901)
       - Quantiles: 
          ╔══════════════════════════════════════════╗
          ║  25%   50%   75%   80%   90%   95%   99% ║
          ║5.330 5.505 5.698 5.735 5.830 5.903 6.013 ║
          ╚══════════════════════════════════════════╝
    - Likelihood ratio: 0.05096 (boot) / 0 (null) = Extremely large 
\end{verbatim}
